---
title: "Alternative TSS usage and TF regulation"
author:
  - "MOCHA authors"
date: "`r BiocStyle::doc_date()`"
package: MOCHA
output:
  BiocStyle::html_document:
    toc_depth: 2
    toc_float: true
vignette: >
  %\VignetteIndexEntry{Alternative TSS usage and TF regulation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  eval = FALSE
)
tutorial_read_chunks <- function(...) {
  files <- unlist(list(...))
  for (f in files) {
    p <- system.file("tutorials", f, package = "MOCHA")
    if (!nzchar(p)) p <- file.path("..", "inst", "tutorials", f)
    knitr::read_chunk(p)
  }
}
tutorial_read_chunks("06-alt-tss-motifs.R")
```

## Introduction

After differential testing, complete the [MOCHA workflow tutorial](MOCHA-workflow-tutorial.html)
through `getDifferentialAccessibleTiles()` so you have `sampleTileMatrices` and
`differentials` (`GRanges` with `FDR` and `Log2FC_C`). MOCHA then offers three downstream-analysis
modules that correspond to Figures 4–5 of the manuscript:

1. **Motif enrichment** within differentially accessible regions.
2. **Alternative TSS regulation** — identifying genes whose multiple TSSs
   respond differently across conditions.
3. **Motif footprinting** — average insertion profiles around TF motif
   instances within open chromatin.

All chunks in this vignette are shown with `eval = FALSE` because the
bundled example data is intentionally minimal. Code is maintained in
`inst/tutorials/06-alt-tss-motifs.R`.

## Setup

```{r libraries, eval=TRUE}
```

We assume you already have:

* `sampleTileMatrices` — a `RangedSummarizedExperiment` from
  `getSampleTileMatrix()`.
* `differentials` — a `GRanges` from
  `getDifferentialAccessibleTiles()` with `FDR` and `Log2FC_C` columns.
* TxDb / OrgDb references registered in your MOCHA object metadata.

## Motif annotation with `addMotifSet()`

Before enrichment or footprinting, attach PWM positions to your sample-tile
matrix (requires `chromVARmotifs`, `motifmatchr`, and a BSgenome):

```{r add-motif-set}
```

## 1. Motif enrichment in differential regions

`MotifEnrichment()` runs a hypergeometric enrichment test for each motif
in a `GRangesList` of motif positions against a background of non-target
regions.

```{r motif-enrichment}
```

`MotifSetEnrichmentAnalysis()` is a higher-level wrapper for upstream TF
regulators:

```{r motifset-enrichment}
```

## 2. Alternative TSS regulation (`getAltTSS`)

```{r get-alt-tss}
```

Locus-level visualization:

```{r plot-alt-tss}
```

## 3. Motif footprinting

```{r motif-footprint}
```

`plotMotifs()` can return summary statistics when `returnDF = TRUE`:

```{r plot-motifs}
```

## Session information

```{r session-info, eval=TRUE}
```
