---
title: "MOCHA quick start with bundled example data"
author:
  - "MOCHA authors"
date: "`r BiocStyle::doc_date()`"
package: MOCHA
output:
  BiocStyle::html_document:
    toc_depth: 2
    toc_float: true
vignette: >
  %\VignetteIndexEntry{MOCHA quick start with bundled example data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE
)
```

## Introduction

MOCHA (Model-based single-cell Open Chromatin Analysis) provides peak calling,
sample-tile matrices, and downstream analyses for scATAC-seq after cell-type
identification. This vignette runs an end-to-end workflow on small bundled
example data (`exampleFragments`, `exampleCellColData`, `exampleBlackList`)
so it can be built on the Bioconductor build system without an ArchR project.

For ArchR-based workflows and larger cohorts, see the package website and
`ImportingFromOtherSources` notebook in `inst/notebooks/`.

## Setup

```{r libraries}
library(MOCHA)
library(MultiAssayExperiment)
library(SummarizedExperiment)
```

Annotation packages used below are listed in `DESCRIPTION` under `Suggests`.

```{r check-deps, echo=FALSE}
has_deps <- all(
  requireNamespace("BSgenome.Hsapiens.UCSC.hg19", quietly = TRUE),
  requireNamespace("TxDb.Hsapiens.UCSC.hg38.knownGene", quietly = TRUE),
  requireNamespace("org.Hs.eg.db", quietly = TRUE)
)
if (!has_deps) {
  stop(
    "Install suggested packages: BSgenome.Hsapiens.UCSC.hg19, ",
    "TxDb.Hsapiens.UCSC.hg38.knownGene, org.Hs.eg.db"
  )
}
```

## Peak calling with `callOpenTiles`

MOCHA accepts a per-sample `GRangesList` of fragments plus cell metadata.

```{r call-open-tiles}
tileResults <- callOpenTiles(
  ATACFragments = exampleFragments,
  cellColData = exampleCellColData,
  blackList = exampleBlackList,
  genome = "BSgenome.Hsapiens.UCSC.hg19",
  TxDb = "TxDb.Hsapiens.UCSC.hg38.knownGene",
  OrgDb = "org.Hs.eg.db",
  outDir = tempdir(),
  cellPopLabel = "Clusters",
  cellPopulations = c("C2", "C5"),
  numCores = 1,
  verbose = TRUE
)
tileResults
```

Inspect called open tiles for one population:

```{r get-open-tiles}
openC2 <- getOpenTiles(tileResults, cellPopulations = "C2")
length(openC2[[1]])
```

## Sample-tile matrices

Consensus tiles across samples are summarized with `getSampleTileMatrix`.

```{r sample-tile-matrix}
sampleTileMatrices <- getSampleTileMatrix(
  tileResults,
  cellPopulations = c("C2", "C5"),
  reproducibilityThreshold = 0
)
sampleTileMatrices
```

Subset to one cell population:

```{r subset}
c2Matrix <- subsetMOCHAObject(
  sampleTileMatrices,
  subsetBy = "celltype",
  groupList = "C2"
)
dim(c2Matrix)
```

## Annotate tiles

Gene annotations are added from TxDb/OrgDb stored in object metadata.

```{r annotate}
annotated <- annotateTiles(c2Matrix)
head(SummarizedExperiment::rowData(annotated))
```

## Differential accessibility (illustration)

When sample metadata includes a grouping column, differential testing compares
foreground and background sample sets. The example data uses the `Sample`
column; we compare the first two samples as a minimal demonstration.

```{r differentials}
samples <- unique(SummarizedExperiment::colData(c2Matrix)$Sample)
if (length(samples) >= 2L) {
  diffs <- getDifferentialAccessibleTiles(
    SampleTileObj = c2Matrix,
    cellPopulations = "C2",
    groupColumn = "Sample",
    foreground = samples[1],
    background = samples[2],
    numCores = 1,
    verbose = TRUE
  )
  head(diffs)
}
```

## Session information

```{r session-info}
sessionInfo()
```
