---
title: "MOCHA workflow tutorial"
author:
  - "MOCHA authors"
date: "`r BiocStyle::doc_date()`"
package: MOCHA
output:
  BiocStyle::html_document:
    toc_depth: 2
    toc_float: true
vignette: >
  %\VignetteIndexEntry{MOCHA workflow tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE
)
vignette_setup <- system.file("tutorials", "vignette_setup.R", package = "MOCHA")
if (!nzchar(vignette_setup)) {
  vignette_setup <- file.path("..", "inst", "tutorials", "vignette_setup.R")
}
source(vignette_setup, local = FALSE)
vignette_read_tutorial("01-workflow.R")
```

## Introduction

MOCHA (Model-based single-cell Open Chromatin Analysis) provides peak calling,
sample-tile matrices, and downstream analyses for scATAC-seq after cell-type
identification. This vignette runs an end-to-end workflow on small bundled
example data (`exampleFragments`, `exampleCellColData`, `exampleBlackList`)
so it can be built on the Bioconductor build system without an ArchR project.

Runnable code is maintained in `inst/tutorials/01-workflow.R` (included here via
`knitr::read_chunk()`).

For importing from ArchR, Signac, SnapATAC2, and other sources and calling open
tiles, see [Data Import Tutorial: Signac, ArchR, SnapATAC and more](Data-Import-Tutorial.html).

## Setup

```{r libraries}
```

Annotation packages used below are listed in `DESCRIPTION` under `Suggests`.

```{r check-deps, echo=FALSE}
```

## Peak calling with `callOpenTiles`

MOCHA accepts a per-sample `GRangesList` of fragments plus cell metadata.

```{r call-open-tiles}
```

Inspect called open tiles for one population:

```{r get-open-tiles}
```

## Tuning parameters

Before building sample-tile matrices, inspect how peak reproducibility
varies across samples. `plotConsensus()` and `suggestConsensusThreshold()`
help you choose `reproducibilityThreshold` for `getSampleTileMatrix()`.

The bundled PBMCSmall example has one sample per cell population, so
`suggestConsensusThreshold()` may return an empty table; use
`reproducibilityThreshold = 0` (union) below when that happens.

```{r plot-consensus}
```

```{r suggest-threshold}
```

The table below maps common tuning knobs to the functions that help you
inspect them. For dropout-aware differentials, see
[MOCHA downstream workflows](MOCHA-downstream-workflows.html).

| Knob | Where set | Helper to inspect |
|------|-----------|-------------------|
| `reproducibilityThreshold` | `getSampleTileMatrix()` | `plotConsensus()`, `suggestConsensusThreshold()` |
| `signalThreshold` | tile filtering in STM pipeline | `plotIntensityDistribution()` |
| `method` / `pairColumn` | `getDifferentialAccessibleTiles()` | — |
| `dropoutAdjustment` | `getDifferentialAccessibleTiles()` | `assessDropout()` |

## Sample-tile matrices

Consensus tiles across samples are summarized with `getSampleTileMatrix()`.
We use `reproducibilityThreshold = 0` here for a permissive union; in practice
you may use the value from `suggestConsensusThreshold()` above when multiple
samples are available.

```{r sample-tile-matrix}
```

```{r sample-tile-matrix-tuned, eval=FALSE}
```

Subset to one cell population:

```{r subset}
```

```{r plot-intensity, eval=FALSE}
```

## Annotate tiles

Gene annotations are added from TxDb/OrgDb stored in object metadata.

```{r annotate}
```

### Motif annotations (optional)

To attach PWM motif positions to a sample-tile matrix for enrichment or
footprinting, use `addMotifSet()` (requires `chromVARmotifs`, `motifmatchr`,
and a BSgenome). See [Alternative TSS usage and TF regulation](Alternative-TSS-TF-regulation.html).

```{r add-motif-set, eval=FALSE}
```

## Differential accessibility (illustration)

When sample metadata includes a grouping column, differential testing compares
foreground and background sample sets. The example data uses the `Sample`
column; we compare the first two samples as a minimal demonstration.

```{r differentials}
```

### Advanced differential options

`getDifferentialAccessibleTiles()` supports additional arguments for
paired designs and dropout handling:

```{r differentials-advanced, eval=FALSE}
```

When `dropoutAdjustment` is set and dropout probabilities are not already
stored on the object, MOCHA runs `assessDropout()` automatically.

## Next steps

- [MOCHA downstream workflows](MOCHA-downstream-workflows.html) — co-accessibility,
  dropout modeling, pseudobulk exploration, merging studies.
- [Alternative TSS usage and TF regulation](Alternative-TSS-TF-regulation.html) —
  motif enrichment, alternative TSSs, footprinting.
- [Exporting and sharing MOCHA results](MOCHA-export-and-sharing.html) —
  `packMOCHA()`, coverage export, IGV-oriented outputs.
- [Advanced modeling](MOCHA-advanced-modeling.html) — mixed models and custom
  peak training with `trainPeakModel()`.

## Session information

```{r session-info}
```
