---
title: "Advanced modeling: mixed models and custom peak training"
author:
  - "MOCHA authors"
date: "`r BiocStyle::doc_date()`"
package: MOCHA
output:
  BiocStyle::html_document:
    toc_depth: 2
    toc_float: true
vignette: >
  %\VignetteIndexEntry{Advanced modeling: mixed models and custom peak training}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  eval = FALSE
)
vignette_setup <- system.file("tutorials", "vignette_setup.R", package = "MOCHA")
if (!nzchar(vignette_setup)) {
  vignette_setup <- file.path("..", "inst", "tutorials", "vignette_setup.R")
}
source(vignette_setup, local = FALSE)
vignette_read_tutorial("05-advanced-modeling.R")
```

## Introduction

Beyond Wilcoxon-based differentials in the
[MOCHA workflow tutorial](MOCHA-workflow-tutorial.html), MOCHA supports linear
mixed-effects models (LMEM), zero-inflated generalized linear mixed models
(ZIGLMM), and custom peak training with `trainPeakModel()`.

This vignette is entirely reference-only (`eval = FALSE`) because fitting
requires study-specific metadata, sufficient sample counts, and optional
packages (`lmerTest`, `glmmTMB`). Code is maintained in
`inst/tutorials/05-advanced-modeling.R`.

```{r libraries, eval=TRUE}
```

**Prerequisites:** A sample-tile matrix from `getSampleTileMatrix()` with
appropriate `colData` columns (e.g. `Sample`, `PassQC`, subject IDs for
random effects).

## Linear mixed-effects models

Pilot a formula on a subset of tiles, then run LMEM across the matrix.

```{r pilot-lmem}
```

```{r run-lmem}
```

## Zero-inflated GLMM

For count data with excess zeros, ZIGLMM extends the mixed-model framework.

```{r pilot-ziglmm}
```

```{r run-ziglmm}
```

`linearModeling()` provides a higher-level interface when you want MOCHA to
select between LMEM and ZIGLMM workflows; see `?linearModeling`.

## Custom peak training

`trainPeakModel()` learns sample-specific peak boundaries from fragments when
you need models beyond the default `callOpenTiles()` peak caller.

```{r train-peak-model}
```

Requires fragment-level data and genome annotation packages (see
[Data Import Tutorial](Data-Import-Tutorial.html)).

## MOCHA S4 subclasses (optional)

By default, MOCHA returns standard Bioconductor types. Opt in with
`returnClass = "mocha"` in constructors, or coerce existing objects:

```{r s4-classes}
```

Subclasses inherit `MultiAssayExperiment` / `RangedSummarizedExperiment`
behavior; `subset()` mirrors `subsetMOCHAObject()`.

## Session information

```{r session-info, eval=TRUE}
```
