---
title: "Exporting and sharing MOCHA results"
author:
  - "MOCHA authors"
date: "`r BiocStyle::doc_date()`"
package: MOCHA
output:
  BiocStyle::html_document:
    toc_depth: 2
    toc_float: true
vignette: >
  %\VignetteIndexEntry{Exporting and sharing MOCHA results}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  eval = FALSE
)
tutorial_read_chunks <- function(...) {
  files <- unlist(list(...))
  for (f in files) {
    p <- system.file("tutorials", f, package = "MOCHA")
    if (!nzchar(p)) p <- file.path("..", "inst", "tutorials", f)
    knitr::read_chunk(p)
  }
}
tutorial_read_chunks("04-export.R")
```

## Introduction

After the [MOCHA workflow tutorial](MOCHA-workflow-tutorial.html) and optional
[downstream workflows](MOCHA-downstream-workflows.html), you may want to share
results with collaborators or visualize tracks in IGV or the MOCHA Dash app.

This vignette ports the narrative from
`inst/notebooks/MOCHA-Export-Vignette.ipynb` into BiocStyle R Markdown. All
chunks are reference-only because exports involve file I/O, large fragment
data, and optional heavy dependencies. Code is maintained in
`inst/tutorials/04-export.R`.

```{r libraries, eval=TRUE}
```

**Prerequisites:** `tileResults` from `callOpenTiles()` and
`sampleTileMatrices` from `getSampleTileMatrix()`.

## Pack and unpack MOCHA objects

`packMOCHA()` archives a MOCHA object and its on-disk outputs into a single zip
for transfer. `unpackMOCHA()` restores the object and rewrites stored paths.

```{r pack-unpack}
```

If you move the project directory after unpacking, update stored paths:

```{r update-path}
```

## Export sample-specific coverage

`exportCoverage()` writes bigWig files from fragment data in a tile-results
object. This step can be slow and disk-intensive; use `numCores` and a
dedicated output directory.

```{r export-coverage}
```

## Export open tiles

```{r export-open-tiles}
```

## Export differential tiles

Load or compute differential peaks (`getDifferentialAccessibleTiles()`), then
export accessibility tracks for foreground/background contrasts.

```{r export-diffs}
```

## Export motifs

Annotate peaks with PWMs (`addMotifSet()`), then export motif positions for
visualization. See also [Alternative TSS usage and TF regulation](Alternative-TSS-TF-regulation.html).

```{r export-motifs}
```

## Export local footprints

```{r export-footprints}
```

## Interactive tracks (Dash app)

For an interactive genome browser built on exported tracks, see
`inst/notebooks/Tracks-Dash-App-Generic.ipynb` in the MOCHA source repository.

## Session information

```{r session-info, eval=TRUE}
```
