---
title: "MOCHA downstream workflows: co-accessibility, dropout, and pseudobulk"
author:
  - "MOCHA authors"
date: "`r BiocStyle::doc_date()`"
package: MOCHA
output:
  BiocStyle::html_document:
    toc_depth: 2
    toc_float: true
vignette: >
  %\VignetteIndexEntry{MOCHA downstream workflows: co-accessibility, dropout, and pseudobulk}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  eval = FALSE
)
vignette_setup <- system.file("tutorials", "vignette_setup.R", package = "MOCHA")
if (!nzchar(vignette_setup)) {
  vignette_setup <- file.path("..", "inst", "tutorials", "vignette_setup.R")
}
source(vignette_setup, local = FALSE)
vignette_read_tutorial("03-downstream.R")
```

## Introduction

This article covers analyses that extend the core workflow in the
[MOCHA workflow tutorial](MOCHA-workflow-tutorial.html): co-accessibility between
tiles, dropout modeling for zero inflation, pseudobulk dimensionality reduction,
merging studies, and object utilities.

**Prerequisites:** `tileResults`, `sampleTileMatrices`, `c2Matrix`, and
`diffs` from the workflow tutorial (or equivalent objects from your project).

All code chunks are reference-only (`eval = FALSE`) so the vignette builds on
Bioconductor without large optional dependencies. Code is maintained in
`inst/tutorials/03-downstream.R` and verified by `tests/scripts/run_tutorials.R`.

```{r libraries, eval=TRUE}
```

## Co-accessibility

Co-accessibility links pairs of open tiles whose accessibility covaries across
samples within a cell population. Start from regions of interest (as `GRanges`
or MOCHA tile strings), then filter and test links. The example uses a synthetic
multisample matrix with a `GroupA` column for contrast testing.

```{r co-access-links}
```

```{r filter-co-access}
```

```{r test-co-access}
```

For chromVAR-based testing on motif-linked regions, see `?testCoAccessibilityChromVar`.

```{r co-access-not-cran, eval=(Sys.getenv("NOT_CRAN") == "true")}
```

## Dropout modeling

Technical zeros in sparse scATAC can bias differential tests. MOCHA estimates
per-tile dropout probabilities and can integrate them into
`getDifferentialAccessibleTiles()` via `dropoutAdjustment`.

```{r assess-dropout}
```

```{r dropout-model}
```

```{r dropout-plots}
```

Use `dropoutAdjustment` in `getDifferentialAccessibleTiles()`; see the workflow
tutorial cheat-sheet and `?getDifferentialAccessibleTiles`.

## Pseudobulk exploration

Summarize sample-level accessibility across tiles for exploratory PCA/UMAP.
Requires `uwot` (Suggests).

```{r bulk-dr}
```

## Combining studies

When you have separate MOCHA runs, merge peak-calling results or combine
sample-tile matrices for joint analysis.

```{r merge-tiles}
```

```{r combine-stm}
```

## Utilities

Common getters and metadata helpers:

```{r getters}
```

```{r rename-add-subset}
```

`StringsToGRanges()` converts MOCHA tile strings to `GRanges` for custom analyses.

## Next steps

- [Exporting and sharing MOCHA results](MOCHA-export-and-sharing.html) for
  `packMOCHA()`, bigWig export, and IGV-oriented outputs.
- [Alternative TSS usage and TF regulation](Alternative-TSS-TF-regulation.html)
  for motif enrichment and footprinting.
- [MOCHA workflow tutorial](MOCHA-workflow-tutorial.html) to revisit peak calling
  and differentials.

## Session information

```{r session-info, eval=TRUE}
```
